---
title: "timeSeries_Demo"

date: "2024-03-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



Some codes and data were borrowed from Dr. Bhaskaran et al
https://github.com/gasparrini/2013_bhaskaran_IJE_codedata/blob/master/londondataset2002_2006.dta

   *Time series regression studies in environmental epidemiology*
   Bhaskaran, Gasparrini, Hajat, Smeeth and Armstrong 
   International Journal of Epidemiology - 2013
   http://www.ag-myresearch.com/2013_bhaskaran_ije.html



Extra lines were inserted to make the codes easier to follow and illustrate the assumptions of time-series data more explicit 


### Install or load libraries

```{r pressure, message=FALSE, warning=FALSE}
# remove all variables to clean the environment 
rm(list = ls())

# First install the package - !! Once installed, remove the installation commande below or comment out by #, as there is no need to install the same lib again and again, all it takes is to use library() command to load installed library.  
#install.packages("foreign")
#install.packages("dlnm")
#install.packages("tsModel")
#install.packages("Epi")
#install.packages("tidyverse")

# laod the package installed if installed already 
library(foreign) 
library(tidyverse)
library(dlnm)
library(Epi)
library(tsModel)
```


## Describe data 
```{r}
# Get file directly from web URL, 
data <- read.dta("https://github.com/gasparrini/2013_bhaskaran_IJE_codedata/raw/master/londondataset2002_2006.dta")

# SET THE DEFAULT ACTION FOR MISSING DATA TO na.exclude
# (MISSING EXCLUDED IN ESTIMATION BUT RE-INSERTED IN PREDICTION/RESIDUALS)
options(na.action="na.exclude")


# Display data, only first 5 lines  
head(data, 5)

# Data structure
str(data)


# SUMMARY
summary(data)


# CORRELATIONS across each variables 
cor(data[,2:4])


# plot correlation - takes bit of time of see this plot 
pairs(data, main = "Pairwise correlation of varaibles")
```




### Make plots for the several time-seriesin the dataset, namely exposure (O3), outcome (death count), temperature, and humidity - the latter two maybe confounders 
Apparently winter is more humid in London
```{r}
# SET THE PLOTTING PARAMETERS FOR THE PLOT (SEE ?par)
# this allow making two pltos in one window
par(mar=c(2, 2, 2, 2), mex=0.8, mfrow=c(2,1))  # Sets smaller margins
oldpar <- par(no.readonly=TRUE)
par(mex=0.8,mfrow=c(4,1))



# SUB-PLOT FOR DAILY DEATHS, WITH VERTICAL LINES DEFINING YEARS
plot(data$date,data$numdeaths,main="Daily deaths over time",
     ylab="Daily number of deaths", xlab = "week")
## add year line vertical 
abline(v=data$date[grep("-01-01",data$date)],col=grey(0.6),lty=1)


# THE SAME FOR OZONE LEVELS
plot(data$date,data$ozone,main="Ozone levels over time",
     ylab="Daily mean ozone level (ug/m3)",xlab="Date")
abline(v=data$date[grep("-01-01",data$date)],col=grey(0.6))
## add year line vertical 
abline(v=data$date[grep("-01-01",data$date)],col=grey(0.6),lty=1)


# Temperature 
plot(data$date,data$temperature,main="Temperature",
     ylab="Daily temp",xlab="Date")
abline(v=data$date[grep("-01-01",data$date)],col=grey(0.6))


# Humidity 
plot(data$date,data$relative_humidity,main="relative humidity",
     ylab="Daily humidity",xlab="Date")
abline(v=data$date[grep("-01-01",data$date)],col=grey(0.6))
```


### Decompose into periodic (seasonal) patterns and a long-term trend, plus random error 
Y axis represents death count       
so, seasonality and downward trends by whatever unobserved factor may confound the association between O3 and neath count   
```{r}
### Plot to decompose time-series 
tsData <- ts(data$numdeaths, frequency = 365)
tsDataDecomposed <- decompose(tsData)
plot(tsDataDecomposed)


#################
# SCALE EXPOSURE
#################
summary(data$ozone)
data$ozone10 <- data$ozone/10

summary(data$ozone10) # double check scaled or not , new variable ozone10 

```

## How much autocorrelation in your outcome??, check with ACF 
Extremely strong across-measurement correlation in the count of death, meaning that count in one day is very similar to the counts in nearby days. ACF stands for autocorrealtion function     
```{r}
acf(data$ozone)
```

# Analysis 

### Lets estimate the association between ozone and death count
### 1. No confounding adjustment, just a crude association between O3 and death count 
We will use a family regression that is designed to capture count data that cannot be negative in value (Poisson regression) 

$$ logE(Y_t) = log(\lambda_t) =  \beta_0 + \beta_{ozone} X_t $$    
$\beta_0$ is intercept term       
$\beta_{ozone}$ is regression coefficient representing the association between ozone and mortality count       


Note that the coefficient below for this analysis is `-0.009364`: negative sign (if exponentiated to convert into the rate ratio interpretation, it is less than 1.0) - ozone is protective of death. Is this a valid result?      
```{r}
################################################################################
# ESTIMATING OZONE-MORTALITY ASSOCIATION
# (CONTROLLING FOR CONFOUNDERS)
############################################

# COMPARE THE RR (AND CI)
# (COMPUTED WITH THE FUNCTION ci.lin IN PACKAGE Epi, TO BE INSTALLED AND LOADED)

# UNADJUSTED MODEL, only including Ozone 
model1 <- glm(numdeaths ~ ozone,data,family=quasipoisson)  
summary(model1)   

```

Rate Ratio indicates the rate of death (# number death count divided by person-time) in response to one unit increase of ozone. The number look deceivingly small, because it is one unit increase (see the range of ozone in y axis)       
See the column for the RR `exp(Est.)`..the ratio is leass than 1.0, so ozone looks protective of dealth, when confounding is not accounted for   
```{r}
ci.lin(model1,subset="ozone",Exp=T)
```


### 2.  Lets control for seasonality as a confounder in our model equation above   
we will use regression techniques to control seasonality buy adding a variable that have wave shape through a data transformation involving sine and cosine function    
```{r}
#####################################
# OPTION 2: PERIODIC FUNCTIONS MODEL
# (FOURIER TERMS)
#####################################
# GENERATE FOURIER TERMS
# (USE FUNCTION harmonic, IN PACKAGE tsModel TO BE INSTALLED AND THEN LOADED)
# 4 SINE-COSINE PAIRS REPRESENTING DIFFERENT HARMONICS WITH PERIOD 1 YEAR

data$time <- seq(nrow(data))
fourier <- harmonic(data$time,nfreq=4,period=365.25)

# FIT A POISSON MODEL FOURIER TERMS + LINEAR TERM FOR TREND
# (USE OF quasipoisson FAMILY FOR SCALING THE STANDARD ERRORS)
model2 <- glm(numdeaths ~ ozone + fourier + time,data,family=quasipoisson)

# COMPUTE PREDICTED NUMBER OF DEATHS FROM THIS MODEL
pred2 <- predict(model2,type="response")

#############
# FIGURE 2B in the manuscript
#############
plot(data$date,data$numdeaths,ylim=c(100,300),col=grey(0.6),
     main="Sine-cosine functions (Fourier terms)",ylab="Daily number of deaths",
     xlab="Date")
lines(data$date,pred2,lwd=2)
```


#### Lets diagnose the model that controlled for seasonality. 
Compare the strength of autocorrelation (undesirable feature to be removed) between this plot and the plot without seasonality correction above. It is not great, but much better than the results from step 1 above ( no control for seasonality) 

```{r}
#############
# Diagnosis of Model 2
#############
acf(residuals(model2), main = "diagnosis of the model that controlled for seasonality")
```



Rate Ratio, rate of death (# number of person-time) in response to one unit increase of ozone 
Now, the RR of ozone is positive, (or greater than 1.0 when exponentiated), meaning that ozone is associated with death after controlling for period effects 

```{r}
ci.lin(model2,subset="ozone",Exp=T)
```






### 3. Now control for temperature too 
```{r}
# (TEMPERATURE MODELLED WITH CATEGORICAL VARIABLES FOR DECILES)
# (MORE SOPHISTICATED APPROACHES ARE AVAILABLE - SEE ARMSTRONG EPIDEMIOLOGY 2006)

cutoffs <- quantile(data$temperature,probs=0:10/10)
data$tempdecile <- cut(data$temperature,breaks=cutoffs,include.lowest=TRUE)

# CONTROLLING FOR TEMPERATURE, plus ozone and seasonality 
model3 <- glm(numdeaths ~ ozone + fourier + time + tempdecile, data,family=quasipoisson)
summary(model3)
```



Lets check RR 
```{r}
ci.lin(model3,subset="ozone",Exp=T)
```


Does not look too much better than step 2, controlling for season only 
...And ACF (autocorrelation) after accounting for temperature
```{r}
acf(residuals(model3))
```











### 4. Now, add lag to exposure: expsoure today will be allowed to impact death count tomorrow, day after tomorrow..and so on using a distributed lag model    
```{r}
# IN PARTICULAR, THE FUNCTION crossbasis PRODUCES THE TRANSFORMATION FOR 
#   SPECIFIC LAG STRUCTURES AND OPTIONALLY FOR NON-LINEARITY
# THE FUNCTION crosspred INSTEAD PREDICTS ESTIMATED EFFECTS

# PRODUCE THE CROSS-BASIS FOR TEMPERATURE
# AS ABOVE, BUT WITH STRATA DEFINED BY INTERNAL CUT-OFFS
cbtempunc <- crossbasis(data$temperature,lag=c(0,7),
                        argvar=list(fun="strata",breaks=cutoffs[2:10]),
                        arglag=list(fun="integer"))

####################################
# CONSTRAINED (LAG-STRATIFIED) DLM
####################################

# PRODUCE A DIFFERENT CROSS-BASIS FOR OZONE
# USE STRATA FOR LAG STRUCTURE, WITH CUT-OFFS DEFINING RIGHT-OPEN INTERVALS 
cbo3constr <- crossbasis(data$ozone,lag=c(0,7),argvar=list(fun="lin"),
                         arglag=list(fun="strata",breaks=c(1,3)))



model4 <- glm(numdeaths ~ cbo3constr + cbtempunc + fourier + date ,data,family=quasipoisson)
pred4 <- crosspred(cbo3constr,model4,at=10)


```


### Lets plot the lagged impact of ozone on mortality counts over days, in RR 
The plot shows the RR of Ozone on day on the same day, next day, day after...so increased ozone actually influence mortality as a carryover effect over days, even after ozone level goes down     
```{r}
#############
# FIGURE 4C
#############

plot(pred4,var=10,type="p",ci="bars",col=1,pch=19,ylim=c(0.99,1.03),
     main="All lag terms modelled together (with costraints)",xlab="Lag (days)",
     ylab="RR and 95%CI per 10ug/m3 ozone increase")

```
Plot autocorrealtion, there is still residual correlation after accounting for all the factors like temperature, humidity, lag      
It still does not look perfect.....some varibale maybe missing        
```{r}
acf(residuals(model4), na.action = na.omit)
```


### Finally, lets assume that death count in previous day is confoudning the count of death today i.e., associated with the exposure ie.g., O3, and death count in next day. In other words, we will control for the lagged outcome of 1 day (previous day) as a confounder. 
```{r}
# RUN THE MODEL AND OBTAIN PREDICTIONS FOR OZONE LEVEL 10ug/m3
data$y_lag <- NA
data$y_lag  <-  lag(data$numdeaths)

model5 <- glm(numdeaths ~ cbo3constr + cbtempunc + fourier + date + y_lag ,data,family=quasipoisson)
pred5 <- crosspred(cbo3constr,model5,at=10)
```




### Now check ACF after controlling for the count of death from previous day   
```{r}
acf(residuals(model5), na.action = na.omit)
```

